The unknown physics of DM sets the rate of SM particle production within the star as well as the initial distribution in space, momentum, and species of the products.
This information is needed to determine if a given DM encounter with a WD results in runaway fusion and with what frequency.
Of course, this can be done precisely for a specific DM model.
In this Section, we describe several general, illustrative classes of DM-WD encounters which demonstrate the explosiveness of ultra-heavy DM interactions.
We also calculate the typical rates at which these events take place in a WD.

\subsection{Classifying DM-WD Encounters}

DM can generically heat the WD medium through the three schematic interactions depicted in Figure \ref{fig:feynman}: DM-DM collisions, DM decays, and DM-SM scattering.
Note that for ultra-heavy DM these can be complicated events involving many (possibly dark) final states, analogous to the interactions of heavy nuclei.
We classify DM candidates into three types according to the interaction that provides the dominant source of heating, and refer to these as collision, decay, and transit candidates.
We additionally make simplifying assumptions about the spatial extent of these interactions.
For collisions and decays, we consider only ``point-like" heating events with all SM products produced in a localized region (smaller than the trigger size).
For transits, we consider only DM-SM scatters that result in a continuous release of particles along the DM trajectory.

We can also classify candidates according to the evolution of the DM itself inside the star.
Generally there will be some loss of DM kinetic energy due to DM-SM scatters---this is either incidental to the eventual heating of the star or represents the dominant heating mechanism.
We consider two simple, limiting cases depending on the magnitude of this energy loss relative to the DM kinetic energy: ``DM wind" and ``DM capture".
In the DM wind scenario, there is negligible energy loss and the DM simply passes through the star.
In the DM capture scenario, the energy loss due to DM-SM scatters is not capable of igniting runaway but is sufficient to stop the DM and cause it to accumulate inside the star.
We consider both scenarios for collision and decay candidates, for which the capture results in a significantly enhanced rate of events.
For simplicity we consider only the wind scenario for transit candidates, though an enhanced explosiveness from slowed DM continuously scattering off stellar constituents is certainly possible in some models.

\subsection{Transits}

\paragraph{Ignition Condition.}
Runaway fusion only occurs in the degenerate WD interior where thermal expansion is suppressed as a cooling mechanism.
The outer layers of the WD, however, are composed of a non-degenerate gas and it is therefore essential that a DM candidate penetrate this layer in order to ignite a SN.
We parameterize this by a DM stopping power $(dE/dx)_\text{SP}$, the kinetic energy lost by the DM per distance traveled in the non-degenerate layer, and demand that
\begin{align}
\label{eq:CrustCondition}
  \left( \frac{d E}{d x} \right)_\text{SP} \ll
  \frac{m_\chi v^2_\text{esc}}{R_\text{env}},
\end{align}
where $R_\text{env} \sim 50 ~\text{km}$ is the width of a WD envelope \cite{KippenhahnWeigert}.

The energy deposited during a continuous heating event such as a DM transit is best described in terms of a linear energy transfer $(dE/dx)_\text{LET}$, the kinetic energy of SM particles produced per distance traveled by the DM.
If these products have a heating length $L_0$ then the relevant energy deposit must at minimum be taken as the energy transferred over the transit distance $L_0$.
Of course, we can always choose to consider energy deposits over a longer segment of the DM trajectory.
Importantly, as per the general condition \eqref{eq:energy_boom_condition} such a deposition is \emph{less} explosive unless $L_0$ is smaller than the trigger size $\lambda_T$.
Thus, we consider the energy deposited in a transit over the larger of these two length scales.
Assuming the energy of the DM is roughly constant over this heating event, the ignition condition for transit heating is:
\begin{align}
\label{eq:transitexplosion}
  \left( \frac{d E}{d x} \right)_\text{LET} \gtrsim
  \frac{\Eboom}{\lambda_T} \cdot \text{max}
  \left\{\frac{L_0}{\lambda_T}, 1 \right\}^2.
\end{align}
Note that the DM stopping power in the non-degenerate layer $(dE/dx)_\text{SP}$ and the linear energy transfer in the degenerate interior $(dE/dx)_\text{LET}$ are possibly controlled by different physics and may have very different numerical values.
In addition, a transit heating event satisfying condition \eqref{eq:CrustCondition} will have negligible energy loss over the parametrically smaller trigger size or heating length $L_0$, validating \eqref{eq:transitexplosion}.

The above argument sums the individual energy deposits along the DM trajectory as though they are all deposited simultaneously.
This is possible if the DM moves sufficiently quickly so that this energy does not diffuse out of the region of interest before the DM has traversed the region.
We therefore require that the diffusion time $\tau_\text{diff} \sim 10^{-12} ~\text{s}$ across a heated region at temperature $T_f$ be larger than the DM crossing-time:
\begin{align}
  \tau_\text{diff} \sim \frac{L^2}{\alpha(T_f)} \gg
  \frac{L}{v_\text{esc}},
\label{eq:SlowDiffusion}
\end{align}
where $\alpha(T)$ is the temperature-dependent diffusivity, and the DM transits at the stellar escape velocity $v_\text{esc} \sim 10^{-2}$.
This condition is more stringent for smaller regions, so we focus on the smallest region of interest, $L = \lambda_T$.
\eqref{eq:SlowDiffusion} is then equivalent to demanding that the escape speed is greater than the conductive speed of the fusion wave front, $v_\text{cond} \sim \alpha(T_f) / \lambda_T$.
Numerical calculations of $v_\text{cond}$ are tabulated in \cite{Woosley}, and indeed condition \eqref{eq:SlowDiffusion} is satisfied for all WD densities.

\paragraph{Event Rate: Wind Scenario.}
The rate of transit events is given by the flux of DM passing through a WD
\begin{align}
  \Gamma_\text{trans} \sim
  \frac{\rho_{\chi}}{m_\chi} R_\text{WD}^2
  \l\frac{v_\text{esc}}{v_\text{halo}}\r^2 v_\text{halo},
\label{eq:TransitFluxCondition}
\end{align}
where $\rho_\chi$ is the DM density in the region of the WD, and $R_\text{WD}$ is the WD radius.
Here $v_\text{halo} \sim 10^{-3}$ is the virial velocity of our galactic halo, and the transit rate contains an $(v_\text{esc}/v_\text{halo})^2 \sim 100$ enhancement due to gravitational focusing.

\subsection{Collisions and Decays}

\paragraph{Ignition Condition.}
For a point-like DM-DM collision or DM decay event releasing particles of heating length $L_0$, ignition will occur if the total energy in SM products satisfies condition~\eqref{eq:energy_boom_condition}.
Such an event will likely result in both SM and dark sector products, so we parameterize the resulting energy in SM particles as a fraction $f_\text{SM}$ of the DM mass.
For non-relativistic DM, the DM mass is the dominant source of energy and therefore $f_\text{SM} \lesssim 1$ regardless of the interaction details.
With this parameterization, a single DM-DM collision or DM decay has an ignition condition:
\begin{equation}
\label{eq:coldecay}
  m_\chi f_\text{SM}  \gtrsim \Eboom \cdot \text{max} \left \{\frac{L_0}{\lambda_T}, 1 \right \}^3.
\end{equation}
We are thus sensitive to DM masses $m_\chi \gtrsim 10^{16} ~\GeV$.

\paragraph{Event Rate: DM Wind.}
DM that is not captured traverses the WD in $R_\text{WD}/v_\text{esc}$, and the rate of DM-DM collisions within the WD parameterized by cross-section $\sigma_{\chi \chi}$ is:
\begin{align}
  \Gamma_\text{ann}
  \sim \l \frac{\rho_\chi}{m_\chi} \r^2 \sigma_{\chi \chi} \l \frac{v_\text{esc}}{v_\text{halo}}\r^3 v_\text{halo} R_\text{WD}^3. 
  \label{eq:collisionDM}
\end{align}
Similarly the net DM decay rate inside the WD parameterized by a lifetime $\tau_\chi$ is:
\begin{align}
 \Gamma_\text{decay}
   \sim \frac{1}{\tau_\chi} \frac{\rho_{\chi}}{m_\chi} \l \frac{v_\text{esc}}{v_\text{halo}}\r R_\text{WD}^3.
  \label{eq:decayDM}
\end{align}

\paragraph{Event Rate: DM Capture.}
For the remainder of this section, all numerical quantities are evaluated assuming a central WD density $n_\text{ion} \sim 10^{31} ~\cm^{-3}$. 
At this density, the relevant WD parameters are approximately: 
\begin{equation}
M_\text{WD} \sim 1.25 ~M_{\astrosun}, ~~~~ R_\text{WD} \sim 4000 ~\text{km}, ~~~~ v_\text{esc} \sim 2 \times 10^{-2}. 
\end{equation}
We also assume a typical WD temperature $T \sim \text{keV}$.

For the DM to be captured in a WD, it must lose energy $\sim m_\chi v^2$, where $v$ is the relative DM velocity (in the rest frame of the WD) asymptotically far away.
Properly, this DM velocity is described by a (boosted) Maxwell distribution peaked at the galactic virial velocity $v_\text{halo} \sim 10^{-3}$. 
Since typically $v \ll v_\text{esc}$, the DM has initial velocity $v_\text{esc}$ in the star and must lose a fraction $(v/v_\text{esc})^2$ of its energy to become captured. 
The physics of DM capture can be made more precise for a specific interaction.
Consider a spin-independent, elastic scattering off ions with cross section $\sigma_{\chi A}$. 
Assuming $m_\text{ion} \ll m_\chi$, the typical momentum transfer in an elastic scatter is $q \sim \mu_{A} v_\text{esc} \sim 200 ~\MeV$, where $\mu_{A} \sim m_\text{ion}$ is the reduced mass of the DM-nuclei system. 
This corresponds to an energy transfer $q^2/m_\text{ion} \sim m_\text{ion} v_\text{esc}^2 \sim 10 ~\MeV$. 
The average number of DM scatters during a full transit of the WD is simply a ratio of the mean free path to the size of the WD
\begin{equation}
\bar{N}_\text{scat} \sim n_\text{ion} \sigma_{\chi A} R_\text{WD}.
\end{equation}
If $\bar{N}_\text{scat} < 1$, then $\bar{N}_\text{scat}$ is the probability for a \emph{single} scatter to occur during the transit. 

The capture of DM in gravitating bodies is very well-studied \cite{Gould}. 
This calculation becomes slightly involved if a single scatter is ever insufficient to capture the DM.
For incoming DM with asymptotic velocity $\sim v_\text{halo}$ (the peak of the Maxwell distribution), multiple scatters are required if
\begin{equation}
B \equiv \frac{m_\text{ion} v_\text{esc}^2}{m_\chi v_\text{halo}^2} \lesssim 1. 
\end{equation}
It can be shown that the general rate of capture is approximately of the form
\begin{equation}
\Gamma_\text{cap} \sim \Gamma_\text{trans} \cdot \text{min}\left \{1, \bar{N}_\text{scat} \text{min}\{B,1\}\right\}.
\end{equation}
Evidently, the assumption that the scatters responsible for slowing the DM are not sufficient to blow up the WD \eqref{eq:transitexplosion} is a valid one for cross sections
\begin{equation}
\sigma_{\chi A} < \l \frac{\Eboom}{\lambda_T} \r \l \frac{1}{m_\text{ion} v_\text{esc}^2} \r \l \frac{1}{n_\text{ion}} \r \sim 10^{-8} ~\cm^2,
\end{equation}
Since the momentum transfer $q \sim 100 ~\text{MeV}$ is of order the inverse nuclear size, it is reasonable to expect the DM coherently scatters off all nucleons in any nucleus. 
The average per-nucleon cross section (spin-independent) is given by
\begin{equation}
\sigma_{\chi A} = A^2 \l \frac{\mu_{A}}{\mu_{n}}\r^2 F^2(q) \sigma_{\chi n} \sim A^4 F^2(q) \sigma_{\chi n},
\end{equation}
where $F^2(q)$ is the Helm form factor \cite{LUX thesis}:
\begin{equation}
F^2(q) = \exp (-q^2 s^2) \left(\frac{3 j_1(q R_n)}{q R_n}\right)^2, ~~~~ s \sim \text{fm}, ~ R_n \sim \sqrt{(A^{1/3})^2 - 5s}. 
\end{equation}
For momentum transfers at a DM velocity $v_\text{esc}$, we find that $F^2(q) \sim 0.1$. 
In general, the form factor should range from $F^2 \sim 10^{-2} - 1$, where for sufficiently $q$ the DM is effectively scattering off individual nucleons. 
We will conservatively take $F^2 \sim 0.1$ for all momentum transfers, and this does not affect the results significantly. 
Currently, the bound on spin-independent DM nuclear elastic scatters from XENON 1T \cite{Xenon} is
\begin{equation}
\label{eq:xenon}
\sigma_{\chi n} < 10^{-45} ~\text{cm}^2 \l \frac{m_\chi}{10^3 ~\GeV} \r.
\end{equation}
Interestingly, any DM candidate whose scattering cross section satisfies the direct detection constraint is necessarily capturing only a fraction $10^{-2}$ or fewer of those DM which transit the star.
Therefore, we are in the regime where capture rate has the following parametric dependence on mass and scattering cross section:
\begin{equation}
\Gamma_\text{cap} \propto \frac{\sigma_{\chi A}}{m_\chi^2}. 
\end{equation}

First we review the evolution of captured DM in the WD ignoring DM collisions or decays. 
Once DM is captured, it eventually thermalizes to an average velocity
\begin{equation}
v_\text{th} \sim \sqrt{\frac{T}{m_\chi}} \sim 10^{-12} \l \frac{m_\chi}{10^{16} ~\GeV}\r^{-1/2}
\end{equation}
and settles at the thermal radius
\begin{align}
R_\text{th} \sim \l \frac{T}{G m_\chi \rho_\text{WD}}\r^{1/2} \sim 0.1 ~\cm \l \frac{m_\chi}{10^{16} ~\GeV}\r^{-1/2},
\end{align}
where its kinetic energy balances against the gravitational potential energy of the (enclosed) WD mass (for simplicity, we take a constant WD density $\rho_\text{WD} \sim n_\text{ion} m_\text{ion}$ within $R_\text{th}$). 
Of course, the timescale for thermalization depends on the nature of the DM-SM interaction.
This has been explicitly calculated in the case that the DM loses energy via elastic nuclear scatters, see \cite{Tinyakov}. 
First, the DM passes through the WD many times before the size of its orbit becomes fully contained within the star.
This occurs after a time
\begin{equation}
t_1 \sim \l \frac{m_\chi}{m_\text{ion}} \r^{3/2} \frac{R_\text{WD}}{v_\text{esc}} \frac{1}{\bar{N}_\text{scat}} \frac{1}{\text{max}\{\bar{N}_\text{scat}, 1\}^{1/2}} \sim 7 \times 10^7~\text{s} \l \frac{m_\chi}{10^{16} ~\GeV} \r^{3/2} \l \frac{\sigma_{\chi A}}{10^{-30} ~\cm^2} \r^{-3/2}. 
\end{equation}
This stage is relevant only if the energy loss after a single transit does not exceed $\sim m_\chi v_\text{esc}^2$
\begin{equation}
\l \frac{m_\text{ion}}{m_\chi} \r \text{max}\{\bar{N}_\text{scat},1\} < 1,
\end{equation}
which is the case for any cross sections which satisfy the XENON bound \eqref{eq:xenon}. 
Subsequently, the DM completes many orbits within the star until dissipation from elastic scatters reduces the orbital size to the thermal radius.
This occurs after a time
\begin{equation}
t_2 \sim \l \frac{m_\chi}{m_\text{ion}} \r \frac{1}{n_\text{ion} \sigma_{\chi A}} \frac{1}{v_\text{ion}} \left \{ 1+\log \l \frac{m_\chi}{m_\text{ion}}\r \right \} \sim 3 \times 10^8~\text{s}\l \frac{m_\chi}{10^{10} ~\GeV} \r \l \frac{\sigma_{\chi A}}{10^{-38} ~\cm^2} \r^{-1}. 
\end{equation}
where $v_\text{ion} \sim \sqrt{\frac{T}{m_\text{ion}}}$.
Note that the time to complete a single orbit is simply the free-fall timescale:
\begin{equation}
\label{eq:freefalltime}
t_\text{ff} \sim \sqrt{\frac{1}{G \rho_\text{WD}}} \sim 0.1 ~\text{s}.
\end{equation}
If the settling time $t_1 + t_2$ is less than the age of the WD, then the DM will begin steadily accumulating at $R_\text{th}$.
However, if the collected mass of DM at the thermal radius ever exceeds the WD mass within this volume, then there is the possibility of self-gravitational collapse of the DM.
The time to collect a critical number of DM particles needed for self-gravitation is given by
\begin{align}
\label{eq:Ncore}
    t_\text{sg} \sim \frac{\rho_\text{WD} R^3_\text{th}}{m_\chi \Gamma_\text{cap}} \sim 10^{4} ~\text{s} \l \frac{m_\chi}{10^{16} ~\GeV} \r^{-1/2} \l \frac{\sigma_{\chi A}}{10^{-30} ~\cm^2} \r^{-1}. .
\end{align}

We now turn towards the rate of DM-DM collisions during various stages of the described evolution. 
After a time $t_1$, the settling DM constitutes a number density throughout the WD volume. 
The total rate of annihilations of this DM population is given by:
\begin{equation}
\Gamma_\text{ann} \sim \Gamma_\text{cap}^2 \l \frac{R_\text{WD}}{R_\text{th}}\r^2 \sigma_{\chi \chi} \frac{m_\chi}{\sigma_{\chi A}} \frac{G}{R_\text{th}^2 v_\text{esc} v_\text{ion}^2}.
\end{equation}
Even if the annihilation rate for infalling DM is too small, a complementary constraint can be derived from DM-DM collisions at the thermal radius. 
As this population of DM only lasts for a time $t_\text{sg}$, the condition that a single collision does not occur before self-gravitation is:
\begin{equation}
\frac{(\Gamma_\text{cap} t_\text{sg})^2}{R_\text{th}^3} \sigma_{\chi \chi} v_\text{th} t_\text{sg} < 1. 
\end{equation}
At this point, the DM sphere will start to shrink.
The ability for a self-gravitating DM sphere to ignite the WD via annihilations is especially interesting. 
Such a process yields novel bounds that also extend to lower DM masses, and will be discussed in detail in forthcoming work \cite{us}. 

Finally, we compute the decay rate of captured DM.  
Of course, this rate is proportional to the number of DM particles within the WD at any given instance.  
In the wind scenario \eqref{eq:decayDM}, this number is given by $\sim (\Gamma_\text{trans} \times t_\text{ff})$.  
In the capture scenario, the number of DM particles available for decay is instead determined by the settling time inside the WD?
\begin{equation}
\Gamma_\text{decay} \sim  \frac{1}{\tau_\chi} \Gamma_\text{cap} t_2
\label{eq:decayDMcap}
\end{equation}
Since, \eqref{eq:decayDMcap} is independent of $\sigma_{\chi A}$ and scales inversely with $m_\chi$, the rate of captured DM decays in the WD is simply enhanced by a factor $\sim \l\frac{v_\text{esc}^3}{v_\text{halo}^2 v_\text{ion}} \r \sim 10^5$ compared to the rate of transiting ``wind" decays \eqref{eq:decayDM}.  